---
title: "Final Project - Citibike route"
#author: "Marco Siqueira Campos"
#date: "2 de dezembro de 2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


####*<span style="color:blue">Marco Siqueira Campos</span>, December 12rd, 2017* 
###DATA 608  
###MSDS CUNY  

This rmd file is part of the final project and describes the preparation, organization and cleaning of the data for the realization of the citibike route graph.
It is fully functional.

The main data, Cibike trip data New York City september 2017 was moved from the link https://s3.amazonaws.com/tripdata/index.html to my Github.

Load package
```{r, warning=FALSE, message=FALSE}
library(lubridate)
```

Read and verify the main data, csv file, from Marco's Github

```{r }
url<-"https://raw.githubusercontent.com/MarcoSCampos/DATA_608/master/201709-citibike.csv"
sep2017<-read.csv(url)
head(sep2017)
```

### Organizing the data 

### Merging station names
```{r }
sep2017$start.end.name <- paste(sep2017$start.station.name, "-", sep2017$end.station.name)

for(i in 1:nrow(sep2017)){
        if(as.character(sep2017$start.station.name[i]) == as.character(sep2017$end.station.name[i])){
                sep2017$start.end.name[i] <- ""
        }
}
```

### Sampling 1/3 of data
```{r }
set.seed(2017)
sep17a <- sep2017[sample(nrow(sep2017), round(nrow(sep2017)/3)), ]
```

### Cleaning - remove noise 
```{r }
sep17a<-subset(sep17a, end.station.latitude != 0)
sep17a<-sep17a[!grepl("Don't Use", sep17a$end.station.name),]
```

### Extracting days of the week and hour 
```{r }
sep17a$startwday<-wday(as.Date(sep17a$starttime, "%Y-%m-%d %H:%M:%S"))
sep17a$starthour<-hour(ymd_hms(sep17a$starttime))
```


### Sub setting with only necessary columns
This remove unnecessary data
```{r }
sep17a <- sep17a[c(16,17,18,6,7,10,11)]
```

### Checking the result
```{r }
head(sep17a)
```

### Save final file as .rda (disable)
The file was saved as .RDA to save space and help to graph download faster.
```{r, eval=FALSE}
saveRDS(sep17a, file="sep17a.Rda")
```


The file is ready to do a graph

